Limited evidence for widespread heterogeneity in psychology

### 2018-06-27 writing/draft-heterogeneity.rmd
Added bibliography and links to OSF

---
title: "Limited evidence for widespread heterogeneity in psychology"
author: "Anton Ohlsson Collentine"
header-includes:
   - \setlength\parindent{24pt}
   - \usepackage{setspace}
   - \doublespacing
output: word_document
---

@@

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables

@@

source("../code/helper_functions_tables_figures.r") #Load functions to prep data for figures
```
Empirical research is typically portrayed as proceeding in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, inferential statistics are used  to minimize the risk that a discovery is due to sampling error.  In the second stage, one may ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [ref: Campbell & Stanley]. This extent is often explored through replications of the original study that are either as similar as possible to the original (called 'direct' or 'exact' replications) or with some deliberate variation on conceptual factors (so-called 'conceptual' or 'indirect' replications) [ref: Zwaan et al], and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. In this paper we examine what degree of heterogeneity can be expected in replication studies in psychology and explore a proposed relationship with effect size. We focus in this paper on direct replications.

@@

 Second, unaccounted for heterogeneity is an indication of incomplete theory, since it suggests that a theory is unable to predict all contextual factors of importance to its claims. As such, heterogeneity might imply previously unknown predictors, so called 'hidden moderators' [ref: van Bavel], the discovery of which can be seen as an opportunity for theoretical advancement [ref: Tacket et al]. The corollary is that, unless a theory has predefined additional assumptions ('auxiliary assumptions' [ref: Earp & Trafimow]) which were not fulfilled in the replications, an absence of heterogeneity should generally preclude referral to hidden moderators.

 Third,  the possibility of heterogeneity can create controversy in the interpretation of single replication results. The proclamation of a  'failure' to replicate an effect often seems to imply that the original find was merely a statistical fluke, due to 'p-hacking' [Ref: Simmons et al] or publication bias [ref.. Inzlicht?]. Unsurprisingly, some researchers take offense [e.g., ref: Baumeister]. An alternative explanation for non-replication, often espoused by the original authors [e.g, IJzerman; Strack], is that the effect is more heterogenous than (perhaps implicitly) claimed originally. Such explanations may be valid or not, but even if valid, an effect is typically of less general interest the more specific circumstances it requires to appear. A way for original authors to align researchers' expectations of heterogeneity is to preemptively specify their own, that is, the degree of heterogeneity that would make they themselves lose interest in an effect (e.g., by declaring 'constraints on generality' [ref Simons et al].

Fourth, meta-analytic techniques are affected by heterogeneity. Techniques that attempt to correct for publication bias in their estimate tend to fail in the presence of heterogeneity [refs: Inzlicht; McShane], which is problematic considering widespread publication bias [refs] and its impact on meta-analytic estimates. In addition, heterogeneity alters the interpretation of meta-analytic estimates as either _the_ true effect size (under homogeneity) or the average of the true effect sizes (under heterogeneity), though one may question the usefulness of interpreting the average true effect size in the presence of heterogeneity [ref: blog Simonsohn], just as it may be questionable to interpret an average main effect in the context of an interaction effect [ref Aiken and West 1993].

It is a commonly believed that heterogeneity is the norm in psychology.  In support of this notion, recent large scale reviews of meta-analyses in psychology [ref: Stanley; van Erp et al] report median heterogeneity levels that can best be described as 'large' (see next paragraph). In comparison, the median heterogeneity estimate in medicine would be considered 'small' by the same standard [ref: Ioannidis  "uncertainty in heterogeneity"]. It may simply be that effects in psychology are more heterogenous than those of medicine. However, meta-analyses in psychology also typically include more studies than those in medicine, and it could be that they tend to include studies from a much broader spectrum than meta-analyses in medicine. The median number of studies (effects) per meta-analysis in the psychology sample of [van Erp et al] was 12, whereas in the medicine sample of  [Ioannidis; uncertainty in heterogeneity] the median number of studies was only 7, despite exclusion of all meta-analyses of 3 or fewer studies. It is difficult to separate these explanations. To facilitate doing so, in this paper we focus on meta-analyses of only direct replications, which are exempt from the potential problem of including too disparate studies. In reference to meta-analyses of direct replications, several authors [ref: McShane et al; Tacket et al] have argued that if we were to expect heterogeneity to be absent or minimal anywhere, it would be in pre-registered multi-lab projects with a common protocol, and that heterogeneity has been reported even under such circumstances [e.g., ref: Many Labs 1] is an indication of widespread heterogeneity in psychology. However, even in the case of multi-lab direct replication projects, studies vary on two contextual factors (sample population and settings) and if we believe an effect is sensitive to changes in these two factors we might still expect to find some heterogeneity

One problem lies in the assessment of heterogeneity and its inherent uncertainty. Heterogeneity is often measured by the I2 index [ref: Higgins & Higgins], which allows comparison of estimates across meta-analyses and has an intuitive interpretation. It can be interpreted as the percentage of variability in effect sizes that is due to heterogeneity amongst the true effects (that is, sensitivity to contextual factors) rather than sampling variance, and ranges from 0-100%. The I2 index is a transformation of the well-known Q-test of heterogeneity in meta-analysis [ref: Cochran] by taking $[Q - (k - 1)] / Q \times 100$%, where k is the number of studies included in the meta-analysis and a negative I2 is set to zero.  [Higgins et al] tentatively defined I2-values of 25, 50, and 75% as small/medium/large heterogeneity respectively. Tests of heterogeneity typically have low statistical power in many practical situations [ref: Jackson; Huedo-Medina]. This complicates the discussion of heterogeneity, because while I2 always provides an estimate of heterogeneity, this estimate is often surrounded by wide confidence intervals [ref: Ioannidis  "uncertainty in heterogeneity"]. For example, Ioannidis reports that in a large set of Cochrane meta-analyses, all meta-analyses with I2 point estimates of 0% had upper 95% confidence intervals that exceeded 33%, exceeding what [Higgins] defined as 'small' heterogeneity. In addition, under homogeneity I2 has a central chisquare distribution [ref: von Hippel], a distribution that is right-skewed with more than 40% of observations falling above the expected value (for all k > 4). In other words, even in the absence of heterogeneity, a meta-analysis of 5 or more studies will have an I2 point estimate above zero in more than 40% of cases. Heterogeneity estimates may thus be congruent with a wide range of true heterogeneity levels. Despite exhortations to the contrary [ref: Ioannidis "uncertainty in heterogeneity"] , it remains common to omit confidence intervals in the reporting of I2. In consideration of such uncertainty and the prevalent belief that heterogeneity is the norm in psychology, we examine the existing evidence for heterogeneity in psychology using a sample of pre-registered multi-lab direct replication projects.

Effect size likely influences heterogeneity. Intuitively, it makes sense to believe that if there is no meta-analytic effect there is nothing to moderate (i.e., no heterogeneity). However, a null or near null effect size estimate may arise from failure to consider contextual factors ('hidden moderators', [ref: van Bavel]) and does not by itself imply the absence of heterogeneity. A large meta-analytic effect size on the other hand, can be expected to be associated with more heterogeneity. To illustrate, consider a meta-analysis of say, the correlation between neuroticism and procrastination. Each included study would need to measure the two variables somehow, possibly the same way across studies. However, because of individual differences, measurement is not equally exact for everyone, and between studies the measurement reliabilities may differ either due to sampling variance (that the sample happens to be more or less homogenous) or to differences in contextual factors (e.g., sampling population, method of measurement). This means that even if the underlying true effect size is the same, the observed correlation between the two variables will differ between studies. Keeping measurement reliabilities constant, differences in observed effect sizes will increase with the underlying true effect, resulting in more variability being ascribed to heterogeneity. More formally, an observed correlation $r_{xy}$ can be expressed as the product of the true correlation or effect size, $\rho_{xy}$, multiplied by the square root of the measurement reliabilities for X ($R_{xx'}$) and Y ($R_{yy'}$): $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$. As such, if $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ stays constant while $\rho_{xy}$ increases the observed differences between studies will also increase, thereby increasing heterogeneity of observed effect sizes (see Table 1). We explore the association between effect size and heterogeneity in a sample of pre-registered multi-lab replication projects.

@@

We examine the evidence for widespread heterogeneity in psychology and explore the correlation between effect size and heterogeneity, using 10 pre-registered multi-lab replication projects in psychology (Table 2). These 10 projects examined a total of 37 primary outcome variables and arguably represent the best, least biased, meta-analytic data currently available in psychology. To better interpret heterogeneity estimates we also estimate power of each project to find zero/small/medium/large heterogeneity as defined by [Higgins et al]. Of the four contextual factors a study might vary on (sample population, settings,  treatment variables, and measurement variables) the studies in these projects vary only on two: sample population and settings. Nonetheless, if we believe an effect is sensitive to changes in these two factors we may still expect to find some heterogeneity. For example, the variety of countries and samples involved in some of the replication projects, not all of which are WEIRD (Western, Educated, Industrial, Rich, Democratic [ref: Weird]) might be a reason to expect heterogeneity. On the other hand, it may be that the less disparate collection of studies included in these direct replication meta-analyses as compared to most psychological meta-analyses result in low heterogeneity estimates, similar to estimates in medicine. Regardless of outcome, the sampled multi-lab replication projects can only inform us of susceptibility to changes in the varied contextual factors  (sample population and settings), with no information on susceptibility to changes in other contextual factors.

@@

@@

All code and data for this project are available on the Open Science Framework (OSF) at osf.io/XXXX. We refer to relevant files on OSF using brackets and links in the sections below, that is (osf.io/XXXX/). We ran all analyses using R version X [ref: R citation].

We downloaded and collated summary data from ten pre-registered multi-lab replication projects in psychology (Table 2). Data from all ten projects were publicly available on the Open Science Framework (osf.io) and downloaded between 2018/02/01 and 2018/03/31. Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/XXXX/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size, and additional information related to each effect (see codebook; osf.io/XXXX/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. Finally, we collated the summary data for all effects into one dataset for analysis (osf.io/XXXX/).

To examine heterogeneity across effects, we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran all analyses as specified by the replication authors (osf.io/XXXX/). All effects were estimated with random-effects models and the Restricted Maximum Likelihood (REML) estimator using the R-package metafor [ref], though with a variety of outcome variables: correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. ML1 transformed odd ratios into standardized mean differences when meta-analyzing under the assumption that responses followed logistic distributions [ref: Sanchez-Meca; metafor]. Two projects (RRR5 and RRR7) used the Knapp and Hartung adjustment of the standard errors [ref: Knapp & Hartung] and ML3 correlations were corrected for bias [ref: Hedges 1989; metafor]. After estimating effects, ML3 transformed correlations into eta-squared for reporting, which we did not. For each effect we estimated $I^2$ and its 95% confidence interval. Confidence intervals were calculated in metafor which uses the Q-profile method [ref Jackson 2014].

>>>>

---
title             : "Limited evidence for widespread heterogeneity in psychology"
shorttitle        : "Limited evidence for heterogeneity"

author: 
  - name          : "Anton Ohlsson Collentine"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Warandelaan 2, 5037 AB Tilburg"
    email         : "j.a.e.olssoncollentine@tilburguniversity.edu"
  - name          : "Jelte Wicherts"
    affiliation   : "1"
    corresponding : no    # Define only one corresponding author
  - name          : "Marcel van Assen"
    affiliation   : "1,2"
    corresponding : no    # Define only one corresponding author

affiliation:
  - id            : "1"
    institution   : "Tilburg University"
  - id            : "2"
    institution   : "Utrecht University"

abstract: |
  Abstract to be written.

keywords          : "heterogeneity, meta-analysis, replication, psychology, many labs, registered replication reports"
wordcount         : "3"

bibliography      : ../bibliography/bibliography-heterogeneity.bib

figsintext        : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : no

lang              : "english"
class             : "man"
output            : papaja::apa6_word
---

@@

```{r load packages, data and source functions}
if(!require(readr)){install.packages("readr")}
if(!require(metafor)){install.packages("metafor")}
if(!require(dplyr)){install.packages("dplyr")}
if(!require(purrr)){install.packages("purrr")}
if(!require(kableExtra)){install.packages("kableExtra")}
if(!require(papaja)){install.packages("papaja")}

library(readr) #To load data
library(dplyr) #For data transformation
library(purrr) #For data iteration
library(metafor) #To run meta-analyses
library(kableExtra) #For creating good-looking pdf tables
library("papaja") #APA formatting of .rmd -> word

@@

source("../code/helper_functions_tables_figures.r") #Load functions to prep data for figures
```
Empirical research is typically portrayed as proceeding in two stages. First, belief in the existence of an effect is established. Second, its generalizability is examined by exploring its boundary conditions. In the first stage, inferential statistics are used  to minimize the risk that a discovery is due to sampling error.  In the second stage, one may ask to what extent the effect depends on a particular choice of four contextual factors; the 1) sample population, 2) settings, 3) treatment variables and 4) measurement variables [@Campbell2015]. This extent is often explored through replications of the original study that are either as similar as possible to the original (called 'direct' or 'exact' replications) or with some deliberate variation on conceptual factors (so-called 'conceptual' or 'indirect' replications) [ref: Zwaan et al], and once sufficient studies have accumulated through meta-analysis. In meta-analysis, the heterogeneity of an effect size (henceforth referred to as heterogeneity) is a measure of an effect's susceptibility to changes in these four factors. An effect strongly dependent on one or more of the four factors, unless controlled for, should exhibit high heterogeneity. In this paper we examine what degree of heterogeneity can be expected in replication studies in psychology and explore a proposed relationship with effect size. We focus in this paper on direct replications.

@@

Second, unaccounted for heterogeneity is an indication of incomplete theory, since it suggests that a theory is unable to predict all contextual factors of importance to its claims. As such, heterogeneity might imply previously unknown predictors, so called 'hidden moderators' [@10.2139/ssrn.2820883], the discovery of which can be seen as an opportunity for theoretical advancement [@Tackett2017]. The corollary is that, unless a theory has predefined additional assumptions ('auxiliary assumptions'; @10.3389/fpsyg.2015.00621) which were not fulfilled in the replications, an absence of heterogeneity should generally preclude referral to hidden moderators.

Third,  the possibility of heterogeneity can create controversy in the interpretation of single replication results. The proclamation of a  'failure' to replicate an effect often seems to imply that the original find was merely a statistical fluke, due to 'p-hacking' [@10.1177/0956797611417632] or publication bias [@10.2139/ssrn.2659409]. Unsurprisingly, some researchers take offense [e.g., @10.1016/j.jesp.2016.02.003]. An alternative explanation for non-replication, often espoused by the original authors [e.g, @10.2139/ssrn.2698810; @10.1177/1745691616674460], is that the effect is more heterogenous than (perhaps implicitly) claimed originally. Such explanations may be valid or not, but even if valid, an effect is typically of less general interest the more specific circumstances it requires to appear. A way for original authors to align researchers' expectations of heterogeneity is to preemptively specify their own, that is, the degree of heterogeneity that would make they themselves lose interest in an effect (e.g., by declaring 'constraints on generality' [@10.1177/1745691617708630].

Fourth, meta-analytic techniques are affected by heterogeneity. Techniques that attempt to correct for publication bias in their estimate tend to fail in the presence of heterogeneity [@@10.2139/ssrn.2659409; @10.1177/1745691616662243], which is problematic considering widespread publication bias [refs] and its impact on meta-analytic estimates. In addition, heterogeneity alters the interpretation of meta-analytic estimates as either _the_ true effect size (under homogeneity) or the average of the true effect sizes (under heterogeneity), though one may question the usefulness of interpreting the average true effect size in the presence of heterogeneity [@Simonsohn2017], just as it may be questionable to interpret an average main effect in the context of an interaction effect [@Aiken1991].

It is a commonly believed that heterogeneity is the norm in psychology.  In support of this notion, recent large scale reviews of meta-analyses in psychology [@Stanley2017; @10.5334/jopd.33a] report median heterogeneity levels that can best be described as 'large' (see next paragraph). In comparison, the median heterogeneity estimate in medicine would be considered 'small' by the same standard [@10.1136/bmj.39343.408449.80]. It may simply be that effects in psychology are more heterogenous than those of medicine. However, meta-analyses in psychology also typically include more studies than those in medicine, and it could be that they tend to include studies from a much broader spectrum than meta-analyses in medicine. The median number of studies (effects) per meta-analysis in the psychology sample of @10.5334/jopd.33a was 12, whereas in the medicine sample of  @10.1136/bmj.39343.408449.80 the median number of studies was only 7, despite exclusion of all meta-analyses of 3 or fewer studies. It is difficult to separate these explanations. To facilitate doing so, in this paper we focus on meta-analyses of only direct replications, which are exempt from the potential problem of including too disparate studies. In reference to meta-analyses of direct replications, several authors [@10.1177/1745691616662243; @Tackett2017] have argued that if we were to expect heterogeneity to be absent or minimal anywhere, it would be in pre-registered multi-lab projects with a common protocol, and that heterogeneity has been reported even under such circumstances [e.g., @10.1027/1864-9335/a000178 is an indication of widespread heterogeneity in psychology. However, even in the case of multi-lab direct replication projects, studies vary on two contextual factors (sample population and settings) and if we believe an effect is sensitive to changes in these two factors we might still expect to find some heterogeneity

One problem lies in the assessment of heterogeneity and its inherent uncertainty. Heterogeneity is often measured by the $I^2$ index [@10.1002/sim.1186;  @10.1136/bmj.327.7414.557], which allows comparison of estimates across meta-analyses and has an intuitive interpretation. It can be interpreted as the percentage of variability in effect sizes that is due to heterogeneity amongst the true effects (that is, sensitivity to contextual factors) rather than sampling variance, and ranges from 0-100%. The $I^2$ index is a transformation of the well-known Q-test of heterogeneity in meta-analysis [@10.2307/3001666] by taking $[Q - (k - 1)] / Q \times 100$%, where k is the number of studies included in the meta-analysis and a negative $I^2$ is set to zero. @10.1136/bmj.327.7414.557 tentatively defined $I^2$-values of 25, 50, and 75% as small/medium/large heterogeneity respectively. Tests of heterogeneity typically have low statistical power in many practical situations [@10.1002/sim.2481; @10.1037/1082-989X.11.2.193]. This complicates the discussion of heterogeneity, because while $I^2$ always provides an estimate of heterogeneity, this estimate is often surrounded by wide confidence intervals [@10.1136/bmj.39343.408449.80"]. For example, Ioannidis reports that in a large set of Cochrane meta-analyses, all meta-analyses with $I^2$ point estimates of 0% had upper 95% confidence intervals that exceeded 33%, exceeding what @10.1136/bmj.327.7414.557 defined as 'small' heterogeneity. In addition, under homogeneity $I^2$ has a central chisquare distribution [@10.1186/s12874-015-0024-z], a distribution that is right-skewed with more than 40% of observations falling above the expected value (for all k > 4). In other words, even in the absence of heterogeneity, a meta-analysis of 5 or more studies will have an $I^2$ point estimate above zero in more than 40% of cases. Heterogeneity estimates may thus be congruent with a wide range of true heterogeneity levels. Despite exhortations to the contrary [@10.1136/bmj.39343.408449.80], it remains common to omit confidence intervals in the reporting of $I^2$. In consideration of such uncertainty and the prevalent belief that heterogeneity is the norm in psychology, we examine the existing evidence for heterogeneity in psychology using a sample of pre-registered multi-lab direct replication projects.

Effect size likely influences heterogeneity. Intuitively, it makes sense to believe that if there is no meta-analytic effect there is nothing to moderate (i.e., no heterogeneity). However, a null or near null effect size estimate may arise from failure to consider contextual factors ('hidden moderators'; @10.2139/ssrn.2820883) and does not by itself imply the absence of heterogeneity. A large meta-analytic effect size on the other hand, can be expected to be associated with more heterogeneity. To illustrate, consider a meta-analysis of say, the correlation between neuroticism and procrastination. Each included study would need to measure the two variables somehow, possibly the same way across studies. However, because of individual differences, measurement is not equally exact for everyone, and between studies the measurement reliabilities may differ either due to sampling variance (that the sample happens to be more or less homogenous) or to differences in contextual factors (e.g., sampling population, method of measurement). This means that even if the underlying true effect size is the same, the observed correlation between the two variables will differ between studies. Keeping measurement reliabilities constant, differences in observed effect sizes will increase with the underlying true effect, resulting in more variability being ascribed to heterogeneity. More formally, an observed correlation $r_{xy}$ can be expressed as the product of the true correlation or effect size, $\rho_{xy}$, multiplied by the square root of the measurement reliabilities for X ($R_{xx'}$) and Y ($R_{yy'}$): $r_{xy} = \rho_{xy} \times \sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$. As such, if $\sqrt{R_{xx'}} \times \sqrt{R_{yy'}}$ stays constant while $\rho_{xy}$ increases the observed differences between studies will also increase, thereby increasing heterogeneity of observed effect sizes (see Table 1). We explore the association between effect size and heterogeneity in a sample of pre-registered multi-lab replication projects.

@@

We examine the evidence for widespread heterogeneity in psychology and explore the correlation between effect size and heterogeneity, using ten pre-registered multi-lab replication projects in psychology (Table 2). These ten projects examined a total of 37 primary outcome variables and arguably represent the best, least biased, meta-analytic data currently available in psychology. To better interpret heterogeneity estimates we also estimate power of each project to find zero/small/medium/large heterogeneity as defined by @10.1136/bmj.327.7414.557. Of the four contextual factors a study might vary on (sample population, settings,  treatment variables, and measurement variables) the studies in these projects vary only on two: sample population and settings. Nonetheless, if we believe an effect is sensitive to changes in these two factors we may still expect to find some heterogeneity. For example, the variety of countries and samples involved in some of the replication projects, not all of which are WEIRD (Western, Educated, Industrial, Rich, Democratic [@10.1017/S0140525X0999152X]) might be a reason to expect heterogeneity. On the other hand, it may be that the less disparate collection of studies included in these direct replication meta-analyses as compared to most psychological meta-analyses result in low heterogeneity estimates, similar to estimates in medicine. Regardless of outcome, the sampled multi-lab replication projects can only inform us of susceptibility to changes in the varied contextual factors  (sample population and settings), with no information on susceptibility to changes in other contextual factors.

@@

```{r data-summary-table}
#library(dplyr)
#library(kableExtra)

#Shortened version of APA reference
papers <- c('Klein, R. A., Ratliff, K. A., Vianello, M., Adams, R. B., Jr., Bahník, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project.', 
'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication.',
'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication.',
'Alogna, V. K., Attaya, M. K., Aucoin, P., Bahník, S., Birch, S., Birt, A. R., ... & Buswell, K. (2014). Registered replication report: Schooler and engstler-schooler (1990).',
'Eerland, A., Sherrill, A. M., Magliano, J. P., Zwaan, R. A., Arnal, J. D., Aucoin, P., ... & Crocker, C. (2016). Registered replication report: Hart & Albarracín (2011).',
'Hagger, M. S., Chatzisarantis, N. L., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., ... & Calvillo, D. P. (2016). A multilab preregistered replication of the ego-depletion effect.',
'Cheung, I., Campbell, L., LeBel, E. P., Ackerman, R. A., Aykutoglu, B., Bahník, S., ... & Carcedo, R. J. (2016). Registered Replication Report: Study 1 from Finkel, Rusbult, Kumashiro, & Hannon (2002).',
'Wagenmakers, E. J., Beek, T., Dijkhoff, L., Gronau, Q. F., Acosta, A., Adams Jr, R. B., ... & Bulnes, L. C. (2016). Registered Replication Report: Strack, Martin, & Stepper (1988).',
'Bouwmeester, S., Verkoeijen, P. P., Aczel, B., Barbosa, F., Bègue, L., Brañas-Garza, P., ... & Evans, A. M. (2017). Registered Replication Report: Rand, Greene, and Nowak (2012).',
"O'Donnell, M., Nelson, L., McLatchie, N. M., & Lynott, D. J. (2017). Registered Replication Report: Dijksterhuis & van Knippenberg (1998)")

##Full APA reference below
# c('Klein, R. A., Ratliff, K. A., Vianello, M., Adams, R. B., Jr., Bahník, S., Bernstein, M. J., . . . Nosek, B. A. (2014). Investigating variation in replicability: A "many labs" replication project. Social Psychology, 45(3), 142-152.', 
# 'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68-82.',
# 'Ebersole, C. R., Atherton, O. E., Belanger, A. L., Skulborstad, H. M., Allen, J. M., Banks, J. B., ... & Brown, E. R. (2016). Many Labs 3: Evaluating participant pool quality across the academic semester via replication. Journal of Experimental Social Psychology, 67, 68-82.',
# 'Alogna, V. K., Attaya, M. K., Aucoin, P., Bahník, S., Birch, S., Birt, A. R., ... & Buswell, K. (2014). Registered replication report: Schooler and engstler-schooler (1990). Perspectives on Psychological Science, 9(5), 556-578.',
# 'Eerland, A., Sherrill, A. M., Magliano, J. P., Zwaan, R. A., Arnal, J. D., Aucoin, P., ... & Crocker, C. (2016). Registered replication report: Hart & Albarracín (2011). Perspectives on Psychological Science, 11(1), 158-171.',
# 'Hagger, M. S., Chatzisarantis, N. L., Alberts, H., Anggono, C. O., Batailler, C., Birt, A. R., ... & Calvillo, D. P. (2016). A multilab preregistered replication of the ego-depletion effect. Perspectives on Psychological Science, 11(4), 546-573.',
# 'Cheung, I., Campbell, L., LeBel, E. P., Ackerman, R. A., Aykutoglu, B., Bahník, S., ... & Carcedo, R. J. (2016). Registered Replication Report: Study 1 from Finkel, Rusbult, Kumashiro, & Hannon (2002). Perspectives on Psychological Science, 11(5), 750-764.',
# 'Wagenmakers, E. J., Beek, T., Dijkhoff, L., Gronau, Q. F., Acosta, A., Adams Jr, R. B., ... & Bulnes, L. C. (2016). Registered Replication Report: Strack, Martin, & Stepper (1988). Perspectives on Psychological Science, 11(6), 917-928.',
# 'Bouwmeester, S., Verkoeijen, P. P., Aczel, B., Barbosa, F., Bègue, L., Brañas-Garza, P., ... & Evans, A. M. (2017). Registered Replication Report: Rand, Greene, and Nowak (2012). Perspectives on Psychological Science, 12(3), 527-542.',
# "O'Donnell, M., Nelson, L., McLatchie, N. M., & Lynott, D. J. (2017). Registered Replication Report: Dijksterhuis & van Knippenberg (1998): Multilab direct replication of: A variant of study 4 from Dijksterhuis, A., & van Knippenberg, A.(1998). The relation between perception and behavior, or how to win a game of trivial pursuit. Journal of personality and social psychology. Perspectives on Psychological Science."

summary.table <- dat %>% 
  group_by(rp) %>% 
  summarize(k = n_distinct(Site),
            Countries = n_distinct(country), 
            Effects = n_distinct(effect),
            Participants = if(Effects == 1){sum(Ntotal)}
                           else{sum(Ntotal) / Effects}) %>% #average across effects
  mutate(Paper = papers) %>% 
  select(RP = rp, Paper, everything()) #change order of columns

summary.table %>% knitr::kable(digits = 0)
```

**Note.** For studies with several effects the number of participants is the average across effects, rounded to the closest whole number. Participant numbers are those used for primary analyses by original authors (i.e., after exclusions). RP = Replication Project, k = no. primary studies, ML = Many Labs, RRR = Registered Replication Report. (osf.io/kf6pt/)

@@

All code and data for this project are available on the Open Science Framework (OSF) at osf.io/4z3e7/. We refer to relevant files on OSF using brackets and links in the sections below, that is (osf.io/XXXX/). We ran all analyses using R version 3.4.3 [@RCoreTeam2017].

We downloaded and collated summary data from ten pre-registered multi-lab replication projects in psychology (Table 2). Data from all ten projects were publicly available on the Open Science Framework (osf.io) and downloaded between 2018/02/01 and 2018/03/31. Although some projects (e.g. RRR4) reported results from several outcome variables, we only included primary outcome variables as stated in accompanying publications, resulting in a total of 37 effects. For each effect we extracted (osf.io/3bmvc/) summary data (e.g., means and standard deviations) at the level of the lab as specified by the original authors for their primary analysis (i.e., typically after exclusions). We extracted information on the country of each lab, whether participants were physically in the lab for the study, total number of participants per lab, type of effect size, and additional information related to each effect (see codebook; osf.io/uhq4r/). Extracted data were in a variety of formats: Excel (Many labs 1, RRR1 & RRR2), CSV (Many labs 3, RRR3, RRR4, RRR5, RRR6) and as PDF tables (RRR7, RRR8). In two cases (RRR5 and RRR6) it was necessary to download the raw data to extract summary data. Although a particular lab may have participated in several projects, the lab indicator was typically not the same across projects. Even so, we kept the original lab indicators to facilitate comparing observations in our dataset with the original datasets. Finally, we collated the summary data for all effects into one dataset for analysis (osf.io/456fz/).

To examine heterogeneity of each of the 37 effects, we computed meta-analytic estimates for all 37 effects in our dataset (Table 3). We ran all analyses as specified by the replication authors (osf.io/kf6pt/). All effects were estimated with random-effects models and the Restricted Maximum Likelihood (REML) estimator using the R-package metafor [@10.18637/jss.v036.i03], though with a variety of outcome variables: correlations, standardized mean differences, raw mean differences, odds ratios and risk differences. ML1 transformed odd ratios into standardized mean differences when meta-analyzing under the assumption that responses followed logistic distributions [@10.1037/1082-989x.8.4.448; @10.18637/jss.v036.i03]. Two projects (RRR5 and RRR7) used the Knapp and Hartung adjustment of the standard errors [@10.1002/sim.1482] and ML3 correlations were corrected for bias [@10.1037//0021-9010.74.3.469; @10.18637/jss.v036.i03]. After estimating effects, ML3 transformed correlations into eta-squared for reporting, which we did not. For each effect we estimated $I^2$ and its 95% confidence interval. Confidence intervals were calculated in metafor which uses the Q-profile method [@10.1186/1471-2288-14-103].

In order to facilitate interpretation of our results, we estimated type I error and power of the Q-test of heterogeneity for each of the 37 effects under zero/small/medium/large heterogeneity ($I^2$ = 0/25/50/75%  respectively). In addition, we plotted $I^2$-distributions across effects at these four heterogeneity levels and compared them with the distribution of the observed $I^2$ estimates of the 37 effects. To do so we simulated results for each effect given the number of studies, k, and the sample sizes, $N_k$, of those studies (osf.io/gbf4u/). We simulated the distribution of $I^2$ for correlation, standardized mean difference, and mean difference effect size measures. Because the standard error of the observed effect size in case of odds ratios and risk differences is strongly affected by assumptions of the probability of success in both treatment groups, we did not directly simulate the I^2 distribution for these effect size measures. In these cases, we used the information on study sample sizes and treated the effects as mean differences in our simulations.

As our concern was of heterogeneity, for convenience we set the true effect size to zero in our simulations. This does not affect the results for correlations or mean differences as estimates of effect size and heterogeneity for these measures are statistically independent, and was expected to have negligible effect on the results for standardized mean differences because while these estimates of effect size are positively correlated to their standard errors, the within study variance $\sigma^2$ was kept constant across studies. [Marcel/Jelte: Not sure this paragraph is correct, because my grasp of the mathematics is a bit tenuous]. Nonetheless, as a sensitivity test we also ran the analysis with 'medium' effect sizes [@Cohen1988], see Appendix A (osf.io/XXXX/) [appendix not created yet, hence no link].

In case the observed effect was a correlation, one run of a simulation proceeded as follows. First, we randomly sampled K Fisher-transformed true correlations $\rho_k$ from a normal distribution with mean 0 and heterogeneity (variance) $\tau^2$. Fisher's z-transformation is a normalizing correlation transformation that ranges from negative infinity to positive infinity, though except for extreme correlations it stays close to the -1 to 1 range [@10.2307/2331838; Fisher1921]. Second, for each of the K true Fisher-transformed correlations we sampled one Fisher-transformed observed correlation from a normal distribution with mean $\rho_k$ and variance $1 ??? (N_k - 3)$. Finally, we fitted a random-effects meta-analysis with REML and estimated $I^2$ for that run. In the simulations, we varied the between-studies standard deviation $\tau$ between 0.000 and 0.50 in increments of 0.005, and used 1,000 runs at each step to approximate the distribution of $I^2$ at that value for true heterogeneity.
